---
title: "Sok-1005 Assignment 1"
author: "14"
format: pdf
editor: visual
---

## Assignment text. 

Download the four data series, i.e., temperature measurements at the four different locations in the atmosphere; the Lower-Troposphere, Mid-Troposphere, Tropopause, and Lower-Stratosphere. Each dataset is found using its unique URL. 

Use only the Global temperature from each dataset (i.e., the third data column). 

For each of the four locations in the atmosphere, calculate the **12-month(right-aligned) moving average** (or mean) (you can run **?zoo::rollmean** in the R console to understand what the **right-aligned average** means**)** using all the available data points. Also, calculate a new variable called **Average**, which is the average of the four 12-month moving averages. Make a plot of the four-atmosphere locations and their **Average** in the same plot. Note, the original data points are not to be used in the plot. Include labels identifying the location of the temperature measurements, and the **Average**. In the plot, you should only use data from January 1980 and onwards. 

```{r}
#| echo: false
rm(list=ls())
library(tidyverse)
library(zoo)
```

```{r}
#| echo: false
#| warning: false
#| output: false
# making df from the url.  
df_ltrop <- read_table2("https://www.nsstc.uah.edu/data/msu/v6.0/tlt/uahncdc_lt_6.0.txt")

df_midtrop <- read_table2("https://www.nsstc.uah.edu/data/msu/v6.0/tmt/uahncdc_mt_6.0.txt")

df_trop <- read_table2("https://www.nsstc.uah.edu/data/msu/v6.0/ttp/uahncdc_tp_6.0.txt")

df_lstrat <- read_table2("https://www.nsstc.uah.edu/data/msu/v6.0/tls/uahncdc_ls_6.0.txt")

```

```{r}
# Removing the values that i dont need. 
df_ltrop <- df_ltrop[1:which(df_ltrop$Year %in% "Year")-1, ]
df_midtrop <- df_midtrop[1:which(df_midtrop$Year %in% "Year")-1, ]
df_trop <- df_trop[1:which(df_trop$Year %in% "Year")-1, ]
df_lstrat <- df_lstrat[1:which(df_lstrat$Year %in% "Year")-1, ]

# Selecting the coloumns im gonna keep. 
df_ltrop <- df_ltrop[c(1,2,3)] 
df_midtrop <- df_midtrop[c(1,2,3)]
df_trop <- df_trop[c(1,2,3)]
df_lstrat <- df_lstrat[c(1,2,3)]

# Changing the names, easier when i cbind the df´s. 
df_ltrop <- df_ltrop %>% 
  rename(lowtrop = Globe, year = Year, mo = Mo)
df_midtrop <- df_midtrop %>%
  rename(midtrop = Globe, year = Year, mo = Mo)
df_trop <- df_trop %>%
  rename(trop = Globe, year = Year, mo = Mo)
df_lstrat <- df_lstrat%>%
  rename(lowstrat = Globe, year = Year, mo = Mo)

# Making a date coloumn. 
df_ltrop <- df_ltrop %>%
  mutate(Date = ymd(paste(df_ltrop$year, df_ltrop$mo, 1, sep="-")))
df_midtrop <- df_midtrop %>%
  mutate(Date = ymd(paste(df_midtrop$year, df_midtrop$mo, 1, sep="-")))
df_trop <- df_trop %>%
  mutate(Date = ymd(paste(df_trop$year, df_trop$mo, 1, sep="-")))
df_lstrat <- df_lstrat %>%
  mutate(Date = ymd(paste(df_lstrat$year, df_lstrat$mo, 1, sep="-"))) 

# Selecting the coloumns im gonna keep. 
df_ltrop <- df_ltrop[c(3,4)] 
df_midtrop <- df_midtrop[c(3,4)]
df_trop <- df_trop[c(3,4)]
df_lstrat <- df_lstrat[c(3,4)]

# Changing from chr to numeric. 
df_ltrop$lowtrop <- as.numeric(df_ltrop$lowtrop)
df_midtrop$midtrop <- as.numeric(df_midtrop$midtrop)
df_trop$trop <- as.numeric(df_trop$trop)
df_lstrat$lowstrat <- as.numeric(df_lstrat$lowstrat)
```

```{r}
# Creating rollmean for the df´s. 
df_ltrop <- df_ltrop %>%
  mutate(avg_lowtrop = rollmean(lowtrop, k = 12, fill = NA, align = 'right'))
df_midtrop <- df_midtrop %>%
  mutate(avg_midtrop = rollmean(midtrop, k = 12, fill = NA, align = 'right'))
df_trop <- df_trop %>%
  mutate(avg_trop = rollmean(trop, k = 12, fill = NA, align = 'right'))
df_lstrat <- df_lstrat %>%
  mutate(avg_lowstrat = rollmean(lowstrat, k = 12, fill = NA, align = 'right'))
```

```{r}
# cbind the df´s.
df <- cbind(df_ltrop, df_midtrop, df_trop, df_lstrat)

# Selecting coloumns to keep. 
df <- df[c(1,2,3,4,6,7,9,10,12)]

# Filter Date from 1980. 
df <- df %>%
  filter(Date >= '1980-01-01') 

# Creating Average coloumn.
df <- df %>% 
  mutate(df, Average = rowMeans(select(df, c(avg_lowtrop, avg_midtrop, avg_trop, avg_lowstrat)), na.rm = TRUE))

```

```{r}
# Making a figure with average. 
df %>% 
  ggplot() +
  geom_line(aes(x = Date, y = avg_lowtrop, col = "Lower-Troposphere")) +
  geom_line(aes(x = Date, y = avg_midtrop, col = "Mid-Troposphere")) +
  geom_line(aes(x = Date, y = avg_trop, col = "Tropopause")) +
  geom_line(aes(x = Date, y = avg_lowstrat, col = "Lower-Stratosphere")) +
  geom_line(aes(x = Date, y = Average, col = "Average"), size = 1) +
  labs(x = "Year",
       y = "Temperature in celsius",
       title = "Plot of temperature over time") +
    theme_minimal()
```

\
